---
title: "Plot dbh vs. biomass by species"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>"
)
set.seed(1)
```

```{r}
# Setup
library(tidyverse)
library(fgeo.biomass)
```

---

The goal is to plot dbh (x) versus biomass (y) by species ([issue](https://github.com/forestgeo/allodb/issues/73)). 

Let's first drop rows with missing `dbh` values as we can't calculate biomass for them.

```{r}
census <- fgeo.biomass::scbi_tree1 %>% 
  filter(!is.na(dbh))
```

Let's find allometric equations in allodb and calculate biomass.

```{r}
species <- fgeo.biomass::scbi_species
census_species <- census %>% 
  add_species(species, site = "SCBI")

census_equations <- allo_find(census_species)
```

Notice the warning that equations couldn't be found. This is in part because, for this site, some species in the data couldn't be matched with a corresponding species in allodb, and/or because the equations available are not adequate for the dbh range the doesn't include the actual dbh values in the data. Later we will fall back to using generic equations but we don't support that feature yet. For now we can't calculate `biomass` for rows containing those species so we will now drop them.

```{r}
# Useless for now
census_equations %>% 
  filter(is.na(equation_id)) %>% 
  select(rowid, site, sp, equation_id)

# Dropping useless rows to continue
census_equations2 <- census_equations %>% 
  filter(!is.na(equation_id))
```

We can now calculate `biomass`.

```{r}
biomass <- allo_evaluate(census_equations2)
biomass
```

We now learn that some equations couldn't be evaluated. The problem now is that we still don't support the ability to calculate biomass from values of `dba` which results in missing `biomass` values. Let's confirm this

```{r}
census_equations_biomass <- census_equations2 %>% right_join(biomass)
census_equations_biomass %>% 
  filter(is.na(biomass)) %>% 
  select(rowid, site, sp, dbh, eqn, biomass)
```

Let's drop the corresponding rows as we can't do much with them.

```{r}
census_equations_biomass2 <- census_equations_biomass %>% filter(!is.na(biomass))
```

Now let's box plot `biomass` by species.

```{r}
census_equations_biomass2 %>% 
  ggplot(aes(sp, biomass)) +
  geom_boxplot() +
  ylab("biomass [kg]") +
  coord_flip()
```

Let's see what `dbh` versus `biomass` looks like now.

```{r, fig.height=14}
census_equations_biomass2 %>% 
  # Convert agb from [Mg] to [kg]
  mutate(agb_kg = agb * 1e3) %>% 
  ggplot(aes(dbh, biomass)) + 
  # Reference based on allometries for tropical trees
  geom_point(aes(y = agb_kg), color = "grey", size = 4) +
  geom_point(aes(y = biomass, color = sp)) +
  ylab("Reference `agb` (grey) and calculated biomass (black) in [kg]") +
  xlab("dbh [mm]") +
  theme(legend.position = "bottom")
```

And now lets facet the plot by species.

```{r, fig.height=14}
census_equations_biomass2 %>% 
  # Convert agb from [Mg] to [kg]
  mutate(agb_kg = agb * 1e3) %>% 
  ggplot(aes(x = dbh)) +
  geom_point(aes(y = agb_kg), size = 1.5, color = "grey") +
  geom_point(aes(y = biomass), size = 1, color = "black") +
  facet_wrap("sp", ncol = 4) +
  ylab("Reference `agb` (grey) and calculated `biomass` (black) in [kg]") +
  xlab("dbh [mm]") +
  theme_bw()
```
